Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-29 demystify.tex
Updates from Overleaf

\section{CV/Resume}
\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}

The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.

\begin{commentbox}[Vu:]
    Unlike a job application, CV is not as important for PhD admission because we do not screen applicants based on their CV.  Moreover, we care and pay attention more about your LOR, SOP, etc.  So do not spend too much time on your CV, just make sure it is easy to skim through and well-organized around research activities and achievements.
\end{commentbox}

@@

After you apply, you might get interviews. The most common case is that a prof. is interested in working with you and wants to chat with you, e.g., to offer RA (\autoref{sec:ra}). In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.

@@

The timeline for interviews varies.  Faculty set up interviews based on their (busy and erratic) schedule. Some try to get interviews done before the holidays, while others do them after the holidays.
Do not be surprised if you get an interview invitation at the last minute. Some profs. are informal and may just email you to chat (e.g., \emph{``could you chat in an hr?''}), while others might schedule a formal interview (e.g., \emph{``can you chat at 2 pm on Friday or 10 am on Monday?''}).

Some schools do not do interviews at all (\autoref{sec:no-interview}). They review applications and make decisions based on them.  If you do not get an interview, it does not mean you're out.

@@

You should treat the interview as an informal chat. Prepare an \emph{``elevator pitche''} about your research experience and interests. You might also want to have a 5-minute presentation about your research. If a prof. asks you to read a paper, do it and be prepared to discuss it. You should also ask if you need to prepare for coding. Finally, the interview gives you an opportunity to ask questions, e.g., about the program and the professor's research.  You should definitely ask as it shows that you're interested (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask).

@@

Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist). You should hear back from most schools by mid-March (though rejection can come out very late).

@@

\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US. This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}. After this date, CS programs can gauge how many slots remain unfilled.

@@

\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}. Thus, rejection letters are often sent out late (e.g., after April 15th or even much later). Not much you can do here. You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks). In short, you just have to be patient and wait (and also beware that some schools do not send out rejection letters at all).

>>>>

\section{CV}
\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}

The CV provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.

Prepare your CV in such a way that allows reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.  Moreover, you don't need to adhere to the 1-page rule of resume in job searching, CV has no such requirement and is often longer.

\begin{commentbox}[Vu:]
    Unlike a job application, CV is not as important for PhD admission because we do not screen applicants based on their CV.  Moreover, we care and pay attention more about your LOR, SOP, etc.  So do not spend too much time on your CV, just make sure it is easy to skim through (\autoref{sec:ievaluate}) and well-organized around research activities and achievements.
\end{commentbox}

@@

After you apply, you \emph{might} get interviews. The most common case is that a prof. is interested in working with you and wants to chat, e.g., to offer RA (\autoref{sec:ra}). In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.

@@

The timeline for interviews varies.  Faculty set up interviews based on their busy and erratic schedule. Some try to get interviews done before the holidays, while others do them after the holidays.
Do not be surprised if you get an interview invitation at the last minute. Some profs. are informal and may just email you to chat (e.g., \emph{``could you chat in an hr?''}), while others might schedule a formal interview (e.g., \emph{``can you chat at 2 pm on Friday or 10 am on Monday?''}).

Some programs \textbf{do not do interviews} at all (\autoref{sec:no-interview}). They review applications and make decisions based on them.  If you do not get an interview, it does not mean you're out.

@@

You should treat the interview as an informal chat. Prepare an \emph{``elevator pitche''} about your research experience and interests. You might also want to have a 5-minute presentation about your research. If a prof. asks you to read a paper, do it and be prepared to discuss it. You should also ask if you need to prepare for coding. Finally, the interview gives you an opportunity to ask questions, e.g., about the program and the professor's research.  You should definitely ask as it shows that you're interested.  See \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask.

@@

Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out e.g., you're on their waitlist. You should hear back from most schools by mid-March, though rejection can come out much later.

@@

\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US. This is a standard deadline for PhD programs due to the Council of Graduate Schools Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{CGS}). After this date, CS programs can gauge how many slots remain unfilled.

@@

\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}. Thus, rejection letters are often sent out late (e.g., after April 15th or even much later). Not much you can do here. You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks). In short, you just have to be patient and wait, and also beware that some schools do not send out rejection letters at all.
